---
title: "`r params$LANG_TLS$LANG[params$LANG_TLS$LABEL == 'joint_report_title']` - `r params$COUNTRY_NAME`"
author: "PAHO/PAHO"
date: "`r params$LANG_TLS$LANG[params$LANG_TLS$LABEL == 'joint_report_date']`: `r format(Sys.Date(), '%d.%m.%Y')`"
output:
  word_document: default
params:
  joint_base_data: NULL
  joint_immunity_data: NULL
  joint_surveillance_data: NULL
  country_shapes: NULL
  COUNTRY_NAME: ""
  LANG: "SPA"
  LANG_TLS: NULL
  YEAR_EVAL: 2024
---

```{r setup, include=FALSE}
Sys.setlocale(locale = "es_ES.UTF-8")
options(knitr.duplicate.label = "allow", scipen = 999)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, out.width = "90%")

library(leaflet)
library(mapview)
library(knitr)
library(tidyverse)
library(sf)
library(janitor)
library(htmltools)

LANG_TLS        <- params$LANG_TLS
COUNTRY_NAME    <- params$COUNTRY_NAME
YEAR_EVAL       <- params$YEAR_EVAL
joint_base_data <- params$joint_base_data
joint_immunity_data     <- params$joint_immunity_data
joint_surveillance_data <- params$joint_surveillance_data
country_shapes  <- params$country_shapes

ZOOM_CONFIG <- "Mozilla/5.0 (compatible; MSIE 10.6; Windows NT 6.1; Trident/5.0)"

lang_label <- function(label) LANG_TLS$LANG[LANG_TLS$LABEL == label]
# Titles shared with the dashboard carry a trailing " - %s" for the admin1
# location, filled in via sprintf() there. The report already appends
# country/year in the caption itself, so strip that placeholder here.
title_no_loc <- function(label) sub(" - %s$", "", lang_label(label))

joint_risk_levels <- c("Low","Medium","High","Very high")
joint_palette <- c(Low="#92d050",Medium="#fec000",High="#e8132b","Very high"="#920000")
joint_level_display <- c(
  "Low"       = lang_label("LR"),
  "Medium"    = lang_label("MR"),
  "High"      = lang_label("HR"),
  "Very high" = lang_label("VHR")
)
joint_palette_display <- setNames(unname(joint_palette), unname(joint_level_display))

translate_level <- function(x) {
  v <- joint_level_display[as.character(x)]
  ifelse(is.na(x) | is.na(v), lang_label("no_data"), unname(v))
}

if (!("ADMIN1_KEY" %in% names(country_shapes))) {
  country_shapes <- country_shapes %>%
    mutate(
      ADMIN1_KEY = toupper(iconv(ADMIN1, to = "ASCII//TRANSLIT")),
      ADMIN2_KEY = toupper(iconv(ADMIN2, to = "ASCII//TRANSLIT"))
    )
}

html_white_bg     <- htmltools::tags$style(".leaflet-container { background: #FFF; }")
html_round_legend <- htmltools::tags$style(".leaflet .legend i{ border-radius: 5px; }")

fig_label   <- lang_label("joint_figure_label")
tbl_label   <- lang_label("joint_table_label")

# ---- helpers ----
make_counts <- function(data, level_col) {
  counts <- data %>%
    mutate(level = as.character(.data[[level_col]])) %>%
    count(level, name = "n")
  tibble(level = joint_risk_levels) %>%
    left_join(counts, by = "level") %>%
    mutate(n = replace_na(n, 0L), level = factor(level, levels = joint_risk_levels))
}

make_joint_map_shot <- function(data, level_col, outfile) {
  shapes <- country_shapes %>%
    left_join(data, by = c("ADMIN1_KEY" = "admin1_key", "ADMIN2_KEY" = "admin2_key")) %>%
    st_as_sf()
  pal <- colorFactor(palette = joint_palette, domain = joint_risk_levels,
                     levels = joint_risk_levels, na.color = "#666666")
  m <- leaflet(options = leafletOptions(attributionControl = FALSE)) %>%
    addPolygons(
      data = shapes,
      fillColor   = ~pal(get(level_col)),
      fillOpacity = 0.75, weight = 1, color = "#333"
    ) %>%
    addLegend(
      colors  = joint_palette[rev(joint_risk_levels)],
      labels  = rev(unname(joint_level_display)),
      opacity = 0.75, position = "topright"
    ) %>%
    htmlwidgets::prependContent(html_white_bg) %>%
    htmlwidgets::prependContent(html_round_legend)
  mapshot(m, file = outfile, useragent = ZOOM_CONFIG)
  invisible(outfile)
}

make_summary_text <- function(counts_tbl) {
  total <- sum(counts_tbl$n)
  if (total == 0) return("")
  get_n <- function(l) { v <- counts_tbl$n[as.character(counts_tbl$level) == l]; if (!length(v)) 0L else v }
  pct   <- function(n) paste0(sprintf("%.1f", n / total * 100), "%")
  n_lr  <- get_n("Low"); n_mr <- get_n("Medium")
  n_hr  <- get_n("High"); n_vhr <- get_n("Very high")
  sprintf(lang_label("joint_summary_text"),
    total, "", n_vhr, pct(n_vhr), n_hr, pct(n_hr), n_mr, pct(n_mr), n_lr, pct(n_lr))
}

make_summary_kable <- function(counts_tbl) {
  row_labels <- c(lang_label("joint_level_lr"), lang_label("joint_level_mr"),
                  lang_label("joint_level_hr"), lang_label("joint_level_vhr"),
                  lang_label("joint_total_label"))
  total <- sum(counts_tbl$n)
  get_n <- function(l) { v <- counts_tbl$n[as.character(counts_tbl$level) == l]; if (!length(v)) 0L else v }
  ns  <- c(get_n("Low"), get_n("Medium"), get_n("High"), get_n("Very high"))
  pct <- if (total > 0) paste0(sprintf("%.1f", ns / total * 100), "%") else rep("0.0%", 4)
  df <- data.frame(cat = row_labels, n = c(ns, total), pct = c(pct, "100.0%"),
                   stringsAsFactors = FALSE)
  colnames(df) <- c(lang_label("joint_summary_col_cat"),
                    lang_label("joint_summary_col_n"),
                    lang_label("joint_summary_col_pct"))
  knitr::kable(df, align = "lrr", booktabs = TRUE)
}
```

\newpage

# `r lang_label("joint_intro_title")`

`r lang_label("joint_intro_body")`

## `r lang_label("joint_scale_title")`

`r lang_label("joint_scale_note")`

| `r lang_label("joint_cat_label")` | `r lang_label("joint_mr_label")` | Polio |
|---|---|---|
| `r lang_label("LR")` | < 25 | < 35 |
| `r lang_label("MR")` | 25–50 | 35–48 |
| `r lang_label("HR")` | 50–74 | 49–60 |
| `r lang_label("VHR")` | 74–100 | 60–112 |

## `r lang_label("joint_matrix_title")`

`r lang_label("joint_matrix_note")`

| `r lang_label("joint_mr_label")` ↓ / Polio → | `r lang_label("LR")` | `r lang_label("MR")` | `r lang_label("HR")` | `r lang_label("VHR")` |
|---|---|---|---|---|
| **`r lang_label("LR")`** | `r lang_label("LR")` | `r lang_label("MR")` | `r lang_label("MR")` | `r lang_label("HR")` |
| **`r lang_label("MR")`** | `r lang_label("MR")` | `r lang_label("MR")` | `r lang_label("HR")` | `r lang_label("HR")` |
| **`r lang_label("HR")`** | `r lang_label("MR")` | `r lang_label("HR")` | `r lang_label("HR")` | `r lang_label("VHR")` |
| **`r lang_label("VHR")`** | `r lang_label("HR")` | `r lang_label("HR")` | `r lang_label("VHR")` | `r lang_label("VHR")` |

\newpage

# `r lang_label("joint_section_general")`

```{r sec1_counts, include=FALSE}
sec1_counts <- make_counts(joint_base_data, "joint_risk_level")
```

`r make_summary_text(sec1_counts)`

#### `r tbl_label` 1a: `r title_no_loc("joint_table_title")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r table_1a}
make_summary_kable(sec1_counts)
```

#### `r fig_label` 1a: `r title_no_loc("joint_map_title")` — `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_1a}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_base_data, "joint_risk_level", f)
knitr::include_graphics(f)
```

#### `r tbl_label` 1b: `r lang_label("joint_table_data_title")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r table_1b}
dat <- joint_base_data %>%
  transmute(
    !!lang_label("joint_table_admin1") := admin1_label,
    !!lang_label("joint_table_admin2") := admin2_label,
    !!lang_label("joint_mr_pts_label") := measles_risk_points,
    !!lang_label("joint_table_measles_level") := translate_level(measles_risk_level),
    !!lang_label("joint_polio_pts_label") := polio_total_score,
    !!lang_label("joint_table_polio_level") := translate_level(polio_risk_level),
    !!lang_label("joint_risk_level_title") := translate_level(joint_risk_level)
  )
knitr::kable(dat, align = "llrrlrl", booktabs = TRUE)
```

\newpage

# `r lang_label("menuitem_inm_pob")`

```{r sec2_setup, include=FALSE}
sec2_available <- !is.null(joint_immunity_data) && nrow(joint_immunity_data) > 0
```

```{r sec2_counts, include=FALSE, eval=sec2_available}
sec2_counts_joint   <- make_counts(joint_immunity_data, "joint_immunity_risk_level")
sec2_counts_measles <- make_counts(joint_immunity_data, "measles_immunity_risk_level")
sec2_counts_polio   <- make_counts(joint_immunity_data, "polio_immunity_risk_level")
```

```{r sec2_text, results="asis", eval=sec2_available}
cat(make_summary_text(sec2_counts_joint))
```

#### `r tbl_label` 2a: `r title_no_loc("joint_immunity_table_title")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r table_2a, eval=sec2_available}
make_summary_kable(sec2_counts_joint)
```

#### `r fig_label` 2a: `r title_no_loc("joint_immunity_map_title")` — `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_2a, eval=sec2_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_immunity_data, "joint_immunity_risk_level", f)
knitr::include_graphics(f)
```

#### `r fig_label` 2b: `r lang_label("menuitem_inm_pob")` — `r lang_label("joint_label_measles")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_2b, eval=sec2_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_immunity_data, "measles_immunity_risk_level", f)
knitr::include_graphics(f)
```

#### `r fig_label` 2c: `r lang_label("menuitem_inm_pob")` — `r lang_label("joint_label_polio")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_2c, eval=sec2_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_immunity_data, "polio_immunity_risk_level", f)
knitr::include_graphics(f)
```

\newpage

# `r lang_label("menuitem_surv_qual")`

```{r sec3_setup, include=FALSE}
sec3_available <- !is.null(joint_surveillance_data) && nrow(joint_surveillance_data) > 0
```

```{r sec3_counts, include=FALSE, eval=sec3_available}
sec3_counts_joint   <- make_counts(joint_surveillance_data, "joint_surveillance_risk_level")
sec3_counts_measles <- make_counts(joint_surveillance_data, "measles_surveillance_risk_level")
sec3_counts_polio   <- make_counts(joint_surveillance_data, "polio_surveillance_risk_level")
```

```{r sec3_text, results="asis", eval=sec3_available}
cat(make_summary_text(sec3_counts_joint))
```

#### `r tbl_label` 3a: `r title_no_loc("joint_surveillance_table_title")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r table_3a, eval=sec3_available}
make_summary_kable(sec3_counts_joint)
```

#### `r fig_label` 3a: `r title_no_loc("joint_surveillance_map_title")` — `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_3a, eval=sec3_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_surveillance_data, "joint_surveillance_risk_level", f)
knitr::include_graphics(f)
```

#### `r fig_label` 3b: `r lang_label("menuitem_surv_qual")` — `r lang_label("joint_label_measles")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_3b, eval=sec3_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_surveillance_data, "measles_surveillance_risk_level", f)
knitr::include_graphics(f)
```

#### `r fig_label` 3c: `r lang_label("menuitem_surv_qual")` — `r lang_label("joint_label_polio")`, `r COUNTRY_NAME`, `r YEAR_EVAL`.
```{r figure_3c, eval=sec3_available}
f <- tempfile(fileext = ".png")
make_joint_map_shot(joint_surveillance_data, "polio_surveillance_risk_level", f)
knitr::include_graphics(f)
```

